---
title: "hw03_gapminder"
output: github_document
author: "Liming Liu"
date: '2018-09-26'
---

## Use dplyr/ggplot2 to manipulate and explore data

```{r}
library(gapminder)
library(knitr)
library(kableExtra)
library(tidyverse)
library(reshape2)
```

#### 1. Get the maximum and minimum of GDP per capital for all continents.

```{r}
GDP <- gapminder %>% 
  group_by(continent) %>% 
  summarise(min = min(gdpPercap), max = max(gdpPercap))
GDP
```

In order to plot two sets of data in one plot, use melt to get a long format
```{r}
gdp2 = melt(GDP, id="continent")
gdp2 %>% 
  ggplot(aes(continent,value, color = variable)) + geom_point() + ylab("GDP percap")
```

#### 2. Look at the spread of GDP per capita within the continents.

```{r}
df1 <- gapminder %>% 
  group_by(continent) %>% 
  summarise(Q25 = quantile(gdpPercap,0.25), Q50 = quantile(gdpPercap, 0.5), Q75 = quantile(gdpPercap, 0.75), standardDeviation = sd(gdpPercap))
df1
```

Get the box plot of the spread

```{r}
gapminder %>% 
  ggplot(aes(continent, gdpPercap)) + scale_y_log10() + geom_boxplot(aes(fill = continent),alpha = 0.5)
```

Get a bar graph of the standard deviation, must use stat="identity" in geom_bar

```{r}
df1 %>% 
  ggplot(aes(continent, standardDeviation, fill = continent)) + geom_bar(stat="identity")
```

#### 3. How is life expectancy changing over time on different continents?

```{r}
df2 <- gapminder %>% 
  group_by(continent, year) %>% 
  summarise(life = mean(lifeExp))
df2
```

get the line plot of the lifeExp trend

```{r}
ggplot(df2, aes(year, life)) + geom_line(aes(group = continent,color = continent))
```

#### 4. Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy – you decide – a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.

Set the benchmark to 50 years old
```{r}
df3 <- gapminder %>% 
  filter(lifeExp < 50) %>% 
  group_by(year, continent) %>%
  summarise(lowLife = n())
df3
```

We can see there is no Oceania in the table above, check if the Oceania lifeExp is all above 50:

```{r}
gapminder %>% 
  filter(continent == "Oceania")
```
So Oceania has no low lifeExp country through these years, the results in the table is correct.

Get the line graph of low lifeExp number in each continent

```{r}
df3 %>% 
  ggplot(aes(year, lowLife)) + geom_line(aes(group = continent, color = continent))
```




